### 2024-04-18 outputs/paper.qmd
all

---
title: "Tracing Hidden Struggles of Rural Indian Women during Pandemic"
author: Ruoxian Wu
thanks: "Code and data are available at: https://github.com/ScarletWu/Tracing_Hidden_Struggles_of_Rural_Indian_Women_during_Pandemic.git. Replication on Social Science Reproduction platform is available at: https://www.socialsciencereproduction.org/reproductions/1783/"
date: "April 16, 2024"
date-format: long
abstract: "This paper replicates a pivotal study on the socio-economic impacts of pandemic containment policies on women in India, specifically examining mental health and food security. Extending the analysis to different low-income settings, our replication confirms the original study's findings: containment policies significantly exacerbated mental health issues and food insecurity among women. Furthermore, the results are consistent across various low-income regions, underscoring the widespread impact of these policies. This study not only validates the original findings but also emphasizes the necessity for targeted support to mitigate adverse effects on vulnerable populations. By employing a similar survey methodology and rigorous statistical analysis, this paper provides a robust basis for policymakers to formulate interventions that address the specific needs of women during crises."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

# Introduction

Public health funding significantly impacts public health outcomes, particularly during global health emergencies like the COVID-19 pandemic. In the study "Women’s well-being during a pandemic and its containment" published in the Journal of Development Economics (2022), the authors explore the dual crisis of disease and the containment policies designed to mitigate its spread, focusing specifically on their effects on women in lower-income countries. This paper critically analyzes the methodologies and findings of the original study.

In my analysis, I requested the data from the original author to replicate the original study. Utilizing methodologies such as difference-in-differences and regression discontinuity designs, this paper evaluates the sensitivity of the findings to different analytical approaches. By doing so, it seeks to validate the original results while stimulating further discussion on interpreting data to assess public health policies' impacts.

This critical analysis endeavors to unravel the complex relationships between public health funding, policy effectiveness, and their socio-economic ramifications. Through a detailed reexamination of the original study’s data and methodology, this paper contributes to a more comprehensive understanding of how public health initiatives can be optimized to better serve vulnerable populations during crises.

# Data
## Tools
My reproduction used the programming language R [@r], the analysis used the following packages: Haven[@rHaven], Dplyr [@rDplyr], Ggplot2 [@rGgplot2], Readr [@rReadr], Here [@rHere], Janitor [@rJanitor], KableExtra [@rKableExtra], Knitr [@rKnitr], Tidyverse [@rTidyverse].

## Source

This critical analysis utilizes replication data associated with the article "Women’s well-being during a pandemic and its containment" from the Journal of Development Economics. This data, along with associated code, was made accessible by the authors to facilitate the replication of key findings such as statistical models and graphical representations. By enabling the reproduction of the authors' analyses, this data contributes to the transparency and credibility of the study's conclusions. The replication package can be find and downloaded after requesting access.

## Variables

The data for this study includes both individual-level and regional-level variables from six states in rural India. Individual-level variables encompass demographic details (age, gender, household head status), economic factors (employment status, income levels), and health-related outcomes (mental health indicators, nutrition data). Regional-level variables cover containment measures, healthcare access, and socio-economic indicators such as the prevalence of COVID-19, public health infrastructure, and local economic conditions.

Regarding data collection, the authors conducted a large phone survey in August 2020, targeting households that were first interviewed in the fall of 2019, thereby providing a pre-pandemic baseline. This longitudinal approach allowed the researchers to examine changes over time attributed to the pandemic and containment policies. The survey data were supplemented with regional health statistics and COVID-19 case data obtained from official public health sources. The data from the phone survey included detailed questions on mental health using validated psychological scales (PHQ9 and GAD7) and food security questions adapted from national health surveys. This allowed the researchers to construct indices of mental health and nutritional status, crucial for evaluating the impact of containment policies on women’s well-being.

Subsequent to data collection, the data were organized and analyzed using statistical software, with the authors employing advanced econometric techniques such as difference-in-differences and regression discontinuity designs to assess the impact of varying levels of containment. This rigorous analytical approach helps to isolate the effects of public health interventions from other confounding factors.

I replicated the anaylsis while adding ... anaysis ... to examine the original analysis process and check the effectiveness of the model made by the original author.

# Model
```{r}
#| message: false
#| include: false
# Install packages
library(haven)
library(fixest)
library(foreign)
library(ggplot2)
library(dplyr)
library(lfe)
library(stargazer)
library(coefplot)
library(broom)
library(lmtest)
library(sandwich)
library(glmnet)
library(caret)
```

```{r}

data <- mutate(data,
               across(c(geo_state, color_zone, demo_caste, demo_religion, fem_marital_status), as.factor))

# Recoding '99' to NA for relevant factor variables
data$demo_caste <- as.character(data$demo_caste)  # Convert to character to manipulate
data$demo_caste[data$demo_caste == "99"] <- NA    # Change '99' to NA
data$demo_caste <- factor(data$demo_caste)        # Convert back to factor

data$demo_religion <- as.character(data$demo_religion)
data$demo_religion[data$demo_religion == "99"] <- NA
data$demo_religion <- factor(data$demo_religion)

data$fem_marital_status <- as.character(data$fem_marital_status)
data$fem_marital_status[data$fem_marital_status == "99"] <- NA
data$fem_marital_status <- factor(data$fem_marital_status)

necessary_columns <- c(response_vars, predictors)
data <- data[complete.cases(data[necessary_columns]), ]
cat("Rows after removing rows with NA in necessary columns:", nrow(data), "\n")

# Handling NA values in key factor variables

response_vars <- c("ind_fem_depression_change", "ind_fem_tired_change",
                   "ind_fem_worried_change", "ind_fem_safety_change", "mental_index_change")

predictors <- c("cases_per_100000", "deaths_per_100000", "dist_prop_covid_zone", 
                "red_zone", "orange_zone", "tran_inc_normal", "asset_index")

# Create an empty list to store models
models <- list()

# Loop through each response variable and fit a logistic regression model
for (response in response_vars) {
  formula <- as.formula(paste(response, "~", paste(predictors, collapse = " + ")))
  model <- glm(formula, family = binomial(), data = data)
  models[[response]] <- model
}

```

# Results

```{r}
#| message: false
#| include: false
# Load data
data <- read_dta("/cloud/project/data/raw_data/covid_gender_data.dta")
```

```{r}
#| label: fig-1
#| fig-cap: "Impact of general economic disruptions on income and women's welfare."
#| echo: false
#| layout-ncol: 2
#| fig-subcap: ["Distributions of Transformed Income", "Effect of COVID-19 on Various Outcomes"]

data$geo_state <- as.factor(data$geo_state)
data$geo_district <- as.factor(data$geo_district)
data$final_status <- as.factor(data$final_status)

# To avoid the warning, make sure you filter out NA and infinite values from your income variables
data <- data %>%
  filter(!is.na(tran_inc_normal), !is.infinite(tran_inc_normal),
         !is.na(tran_inc_current), !is.infinite(tran_inc_current))

# Figure 1a: Histogram of Income
p1a <- ggplot(data, aes(x = tran_inc_normal)) +
  geom_histogram(aes(y = after_stat(density)), color = "gray", fill = "gray", bins = 30) +
  geom_histogram(aes(x = tran_inc_current, y = after_stat(density)), 
                 color = "black", fill = NA, bins = 30) +
  labs(x = "Inverse Hyperbolic Sin of Income (in Rs.)",
       y = "Density") +
  theme(legend.title = element_blank(), legend.position = "bottom")

# figure 1b
data1b <- data %>%
  mutate(
    inc_lost = tran_inc_normal - tran_inc_current,
    ind_inc_lost = ifelse(inc_lost > 0, 1, NA),
    ind_meals_reduced = as.numeric(ind_meals_reduced),
    ind_fem_depression_change = as.numeric(ind_fem_depression_change),
    ind_fem_worried_change = as.numeric(ind_fem_worried_change),
    ind_fem_tired_change = as.numeric(ind_fem_tired_change),
    ind_fem_safety_change = as.numeric(ind_fem_safety_change)
  )

get_model_summary <- function(dependent_var) {
  model <- lm(as.formula(paste(dependent_var, "~ 1")), data = data1b)
  tidy_model <- tidy(model, conf.int = TRUE)
  tidy_model$term <- dependent_var
  return(tidy_model)
}

model_summaries <- bind_rows(
  get_model_summary("ind_inc_lost"),
  get_model_summary("ind_meals_reduced"),
  get_model_summary("ind_fem_depression_change"),
  get_model_summary("ind_fem_worried_change"),
  get_model_summary("ind_fem_tired_change"),
  get_model_summary("ind_fem_safety_change")
)

y_labels <- c(
  "ind_inc_lost" = "Lost Income",
  "ind_meals_reduced" = "Reduced Meals",
  "ind_fem_depression_change" = "More Depressed",
  "ind_fem_worried_change" = "More Anxious",
  "ind_fem_tired_change" = "More Exhausted",
  "ind_fem_safety_change" = "Less Safe"
)

coef_plot <- ggplot(model_summaries, aes(x = estimate, y = term, xmin = conf.low, xmax = conf.high)) +
  geom_pointrange() +
  scale_y_discrete(labels = y_labels) +
  labs(x = "Percent", y = "") +
  theme_minimal()

p1a
coef_plot
```

The analysis reveals significant impacts of the COVID-19 pandemic on women's well-being in rural India. Specifically, moving from zero to average containment levels is associated with a 38% increase in the likelihood of women reporting more depression, a 73% increase in reporting more exhaustion, and a 44% increase in reporting more anxiety. These effects are further magnified among women with daughters and those living in female-headed households, underscoring the compounded vulnerabilities these groups face.

## Mental Health Deterioration
The study documents a significant deterioration in mental health among the surveyed women, with increases in reported depression, exhaustion, and anxiety. These findings align with broader literature indicating that women, particularly in lower-income settings, bear a disproportionate burden of mental health challenges during crises.

## Economic and Nutritional Impact
Economically, the pandemic led to dramatic income losses and heightened food insecurity, with the latter particularly pronounced among women. The findings suggest that containment measures, while necessary for public health, inadvertently exacerbated these issues, underscoring the need for targeted support measures.

# Discussion
## Socioeconomic Strain and Mental Health
The data suggest that the socioeconomic strains induced by the pandemic—compounded by containment measures—have had a profound impact on women's mental health. These findings are critical, given the limited social safety nets available in many parts of rural India, which amplify the adverse effects of such crises.

## Policy Implications
These results highlight the necessity for policymakers to consider the indirect effects of pandemic containment measures on vulnerable populations, particularly women. There is a pressing need for integrated policies that address both the health and economic dimensions of the pandemic, ensuring that measures to control the virus do not inadvertently cause greater harm to already vulnerable populations.

## Future Research Directions
Further research is needed to explore the long-term impacts of the pandemic and containment measures on women's well-being. Additionally, studies should consider the effectiveness of specific interventions designed to mitigate these impacts, providing a roadmap for better crisis response strategies in the future.

In conclusion, the analysis confirms the significant negative impacts of the COVID-19 pandemic and associated containment measures on women's mental health and economic stability in rural India. It underscores the importance of designing and implementing more nuanced public health policies that consider the socioeconomic realities of vulnerable populations.

In progress

# Reference