### 2024-04-18 outputs/datasheet.qmd
citations

---
title: "Datasheet for Indian women's well-being"
author: 
  - Natalie Gaurav Khanna, Corinne Low, Manisha Shah, Sreyashi Sharmin, and Alessandra Voena
thanks: "Code and data are available at: https://datasets.iza.org/dataset/1395/g2lm-lic-women-s-well-being-during-a-pandemic-and-its-containment. "
date: "2 April 2024"
date-format: "D MMMM YYYY"
abstract: "This datasheet provides detailed information about the dataset for Women’s well-being during a pandemic and its containment."
format: pdf
number-sections: true
---
Programs for “Women’s Well-Being During a Pandemic and Its Containment”

What do the instances that comprise the dataset represent?

Each instance in the dataset represents individual responses from a COVID-19 phone survey, capturing detailed socio-economic and health-related data from women heads of households across Northern India.

How many instances are there in total?

The dataset contains over 1,000 instances, each corresponding to a surveyed individual.

What data does each instance consist of?

Each instance includes data on employment status, mental health indicators, nutritional outcomes, and demographic details such as age, education, and household size.

Is there a label or target associated with each instance?

The dataset does not have a specific label or target; it's primarily used for observational and analytical studies on the impacts of the pandemic.

Is any information missing from individual instances?

Some instances may have missing data due to non-responses or skipped questions in the survey process.
Collection Process

How was the data associated with each instance acquired?

The data was collected through telephone interviews conducted by trained personnel from IDInsight, ensuring comprehensive coverage despite lockdown restrictions.

What mechanisms or procedures were used to collect the data?

Standardized questionnaires were administered, with procedures in place to ensure the reliability and consistency of the data collected across different interviewers and respondents.
Preprocessing/Cleaning/Labeling

Was any preprocessing/cleaning/labeling of the data done?

Preprocessing involved cleaning for inconsistencies, handling missing values, and verifying the accuracy of responses. Data was anonymized to protect the identities of the respondents.

# Uses

Has the dataset been used for any tasks already?

The dataset has been extensively used to analyze the impact of COVID-19 containment measures on women's employment, mental health, and nutritional status, resulting in several academic publications and policy reports.

What other tasks could the dataset be used for?

Beyond its current use, the dataset could inform longitudinal studies on long-term impacts of the pandemic, contribute to comparative studies with other regions, and aid in developing targeted interventions for women in similar crises.

# Distribution

Will the dataset be distributed to third parties?

Yes, the dataset is available for non-commercial use to facilitate further research and policy planning, provided proper attribution is given as per the licensing agreement.

# Maintenance

Who will be supporting/hosting/maintaining the dataset?

The dataset will be maintained by the IZA Data Repository.